### 2013-05-28 Publications/2013-TMI/03-introduction.tex
Advancing Introduction (#81)
I re-structured introduction, pushed forward mathematical background
to the II section (methods). Currently I still have to rewrite
the registration stuff, comments of existing literature about
distortion correction evaluation and the final subsection of I. Also some
todo tags have been placed where needed.

\section{Introduction}
\label{sec:introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
%
\IEEEPARstart{D}{iffusion} MRI (dMRI) is a widely used family of \gls{mr} techniques
\citep{sundgren_diffusion_2004} which recently has accounted for a growing
interest in its application to whole-brain structural connectivity analysis.
This emerging field \citep{hagmann_diffusion_2005,sporns_human_2005}, currently 
includes a large amount of imaging techniques for acquisition, processing, 
and analysis specifically tuned for \gls{dwi} data \citep{daducci_connectome_2012}.

The whole-brain connectivity analysis has given rise to some challenges
towards reliable structural information about the neuronal tracts 
from \gls{dwi} \cite{johansen-berg_using_2009,jones_white_2012}. 
Here, we shall address brain tissue segmentation on diffusion space 
and correction of geometrical distortions inherent to the acquisition 
sequence \citep{hagmann_mr_2012}.

In this work, we will refer as brain tissue segmentation to the precise
delineation of the \gls{csf}-\gls{gm} and \gls{gm}-\gls{wm} interface surfaces.
An accurate brain tissue segmentation is required
to filter the fiber bundles obtained with \gls{dwi} tractography. 
This requirement is usually solved in practice by plainly thresholding the 
\gls{fa}, a well-known scalar map derived from \gls{dwi} which depicts 
the isotropy of water diffusion inside the brain. Also,
it is necessary to locate the intersections of fiber bundles and \gls{gm}.
Moreover, a precise location of the \gls{gm}-\gls{wm} surface is also essential 
to finally achieve a consistent parcellisation of the cortex 
to represent the nodes of the output network. This parcellisation 
is generally defined in a high-resolution and better understood structural 
\gls{mri} of the same 
subject (e.g. \gls{t1} and/or \gls{t2} weighted acquisitions). Even though
some efforts have addressed the study of the robustness of tractography with
respect to intra-subject variability \cite{heiervang_between_2006,
wakana_reproducibility_2007}, these results are restricted to certain regions 
of the brain, only. Therefore, robust and precise segmentation methods are 
required in the whole-brain application. The problem is challenging due to
the much lower resolution of \gls{dwi} (typically around 
$2.0\times2.0\times2.0mm^3$) compared to structural \gls{mri}, and the
existence of geometrical distortions.

\gls{dwi} data are usually acquired with \gls{epi} sequences 
as they allow for very fast acquisitions, but they are known to 
suffer from geometrical distortions due to local field inhomogeneities. 
These artifacts happen along the phase-encoding direction, and are most 
appreciable in the front part of the brain for the strong air/tissue 
interface around the frontal sinuses. A number of methodologies have
been developed to correct for the distortion, and are generically named 
as \emph{\gls{epi}-unwarp} techniques
\cite{holland_efficient_2010,hsu_correction_2009,jezzard_characterization_2005,
reber_correction_2005}. These methods usually 
require the extra acquisition of the magnitude and phase of
the field (``field-mapping''), a condition which is not always met. Some other 
methodologies do not make use of the field-mapping, compensating the distortion
with non-linear registration from structural \gls{mri} or other means
\citep{andersson_modeling_2001}. To our knowledge, there exists no study
on the impact of the \gls{epi} distortion on the variability of tractography
results.

@@

We suggest clustering the current methodologies of template-based segmentation 
methods into three groups. The first group typically adds a shape prior term to 
the energy functional of an evolving active contour~\citep{Bresson2006a,Chan2005,
Chen2002,Cremers2006,Gastaud2004,Paragios2003,Vemuri2003a,Yezzi2003a}.
These methods generally have a explicit description of the expected relative boundary 
locations of the object to be delineated, and some even model the statistical deviations
from this average shape. Closely related to this group are atlas-based segmentation
methods \citep{Gorthi2011,Gorthi2009,Pohl2005,Pohl2006,Wang2006}, where the prior 
imposes consistent voxel-based classification of contiguous regions. Here, the 
presence of more structures than one unique \gls{roi} helps aligning the target image 
with the atlas in a hierarchical fashion. Finally, the third group generalizes 
the atlas to actual images, and the contour is to segment simultaneously two 
different target images, related by a spatial transform to be co-estimated~\citep{Wyatt2003,
Yezzi2003}.

>>>>

\section{Introduction}
\label{sec:introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
%
\IEEEPARstart{D}{iffusion} \gls{mri} is a widely used family
of \gls{mri} techniques \citep{sundgren_diffusion_2004} which recently 
has accounted for a growing interest in its application to structural 
connectivity analysis of the brain. This emerging field exploits
\gls{dwi} data to derive the local axonal structure at each imaging voxel 
\citep{basser_microstructural_2011} and estimate a whole-brain mapping of fiber 
tracts represented by trajectories reconstructed from the local information.
This comprehensive map of neural connections of the brain is called the 
\emph{connectome} \citep{hagmann_diffusion_2005,sporns_human_2005}. The connectome
analysis is a promising tool for neuroscience and clinical applications 
(\textbf{FIXME:} why? or references?).

Early \gls{dwi} research focused mainly on the improvements of imaging 
methodologies better understanding the diffusion effect and improving
image reconstruction methodologies. Currently, the connectome extraction 
and analysis relies on a large amount of sophisticated computational techniques
\citep{daducci_connectome_2012,hagmann_mr_2012} including acquisition,
reconstruction, modeling and model fitting, image processing, fibre tracking,
connectivity mapping, visualization, group studies, and inference. This 
growing complexity has given rise to challenging issues towards reliable 
structural information about the neuronal tracts \cite{johansen-berg_using_2009,
jones_white_2012,soares_hitchhikers_2013}, and statistical analysis 
\citep{meskaldji_comparing_????}. Here,we shall address three tasks included within 
the image processing stage in a unified approach: brain tissue segmentation in 
diffusion space (\autoref{sec:dwi_segmentation}), correction of geometrical 
distortions (\autoref{sec:distortion}), and structural image registration to 
diffusion coordinate space (\autoref{sec:registration}). These tasks are generally
solved independently, or combined in pairs. However, there exist fundamental 
coupling relationships that can be exploited to obtain a simultaneous solution to 
the three problems. This joint approach satisfactory impacts the downstream
outcomes of the whole pipeline with the increase of the internal consistency.

\subsection{\gls{dwi} data overview}
\label{sec:dwi_overview}

\todo[inline]{VERY brief look into signal generation HERE and cite reconstruction methods.}

\gls{dwi} data is strongly affected by \glspl{pve} \citep{alexander_analysis_2001},
which appear when several different tissues, or signal emitters, are present
in the same imaging unit, producing an averaged intensity. The effect is directly
related to the low resolution achievable with \gls{dwi} (typically around 
$2.0\times2.0\times2.0mm^3$). An additional complication specific to
\gls{dwi} is the \gls{csf} \emph{contamination} \citep{metzler-baddeley_how_2012},
that is a particular \gls{pve} in which the signal sensed inside the affected voxel is 
linear with respect the \gls{gm} and \gls{wm} contributions, but non-linear with
respect \gls{csf}.

Generally speaking, alongside the difficulty posed by the low resolution, \gls{dwi} 
processing is also challenging due to the \emph{direction dependency} of raw data.
\todo[inline]{Introduce here what are b0, fa, md, and direction dependency problem.}
These low-$b$ (or $b = 0$) volumes are acquired without direction gradient as reference, 
and present a T2-like contrast.

\subsection{\gls{dwi} segmentation}
\label{sec:dwi_segmentation}

A precise delineation of the \gls{csf}, \gls{gm} and \gls{wm} interfacing surfaces
is required with sub-pixel resolution.
The resulting segmentation is necessary to perform the majority of tractography 
algorithms and it is required to filter the resulting tractogram. The \gls{gm}-\gls{wm}
interface is necessary to locate the starting and ending points of the detected
fiber bundles, and \gls{csf}-\gls{wm} surface is critical for pruning spurious 
and discontinued fiber bundles.

A number of methodologies have been proposed for \gls{dwi} segmentation, ranging 
from intensity thresholding to atlas-based segmentation. The first approach is performed 
on the \gls{fa} \citep{ennis_orthogonal_2006}, a well-known scalar map derived from
\gls{dwi} data which depicts the isotropy of water diffusion inside the brain.
Although this methodology was popular among the premier tractography studies,
they were generally limited to certain regions or significant fiber tracts, and thus,
it cannot be applied in whole-brain tractography. Early approaches to \gls{dwi} segmentation 
include level set formulations using scalar maps of direction invariants derived
from the tensor model \citep{zhukov_level_2003}, directly on the diffusion raw data
\citep{rousson_level_2004}, or finding alternative diffusion representations 
\citep{jonasson_representing_2007}. Even though this latter case was restricted to the extraction of
the corpus callosum from a real dataset, the density of the components of the diffusion tensor
are approximated by multivariate Gaussians for first. Iterative clustering performed on the 
\emph{low}-$b$ volumes of \gls{dwi} data was proposed by \citep{hadjiprocopis_unbiased_2005}.
Later studies investigated the application of probabilistic frameworks combining mixtures of 
gaussians, \gls{mrf} and labeling fusion techniques \citep{liu_brain_2007} using as features 
widely-used \gls{dwi}-derived scalar maps as \gls{fa} or \gls{md}. A similar framework 
combining co-registered structural information (T1 weighted) with \emph{independent orthogonal 
invariants} derived from the \gls{dwi} tensor model was proposed by \citep{awate_multivariate_2008}. 
Some proposals suggest the use of the raw diffusion data (directionally dependent) to avoid fitting 
a certain model \citep{lu_segmentation_2008}. In \citep{han_experimental_2009}, graph-cuts 
voxel-based techniques are proposed using the most common diffusion tensor derived features.
Further developments of the probabilistic approach have been proposed adding more scalar maps
as features and a more detailed treatment of \gls{pve} \citep{kumazawa_partial_2010}.

A number of methods have been proposed using features not directly derived from \gls{dwi} data.
Segmentation obtained by co-registering structural T1-weighted images will be covered in 
\autoref{sec:registration}. Some other works delay the segmentation task after the tractogram 
is obtained, performing clustering on features derived on the tracts alignment 
\citep{jonasson_white_2005}, combined tract registration \citep{mayer_supervised_2011} or using
tractography atlases \citep{odonnell_automatic_2007}. However, methods based on tractography
usually address the tractogram segmentation problem, to later combine the solution to 
answer the whole-brain segmentation problem.

None of the presented methods have claimed for definite results, mainly due to the lack of 
a \emph{gold-standard} evaluation methodology. Most of them are tested only on certain
regions \todo{citations}, or do not provide sub-pixel resolution results
\citep{hadjiprocopis_unbiased_2005,liu_brain_2007,awate_multivariate_2008,lu_segmentation_2008,
han_experimental_2009}. Generally, results obtained with high resolution atlas co-registration
(\autoref{sec:registration}) are more compelling, minimizing the activity on this line which is 
currently being considered to be included in reconstruction algorithms 
\citep{kumazawa_improvement_2013}. \emph{Golden}-standard evaluation frameworks have been 
proposed for the segmentation validation \citep{jha_task-based_2012} on the task of lesion 
detection in visceral organs.

\subsection{Correction for susceptibility distortions}
\label{sec:distortion}

\gls{dwi} data are usually acquired with \gls{epi} sequences 
as they allow for very fast acquisitions, but they are known to 
suffer from geometrical distortions and artifacts due to, mainly,
three sources: the subject motion between direction sampling, the 
induced \emph{Eddy currents} on the scanner coils and finally
distortions caused by the magnetic susceptibility inhomogeneity
present at tissue interface. In this paper, we restrict ourselves to
the last one, as it accounts for the major impact in the connectome
analysis. \emph{Susceptibility distortions} happen along the 
phase-encoding direction, and are most appreciable in the front part of 
the brain for the strong air/tissue interface surrounding the frontal sinuses.

One approach to correcting the susceptibility distortions was proposed
with the earliest \gls{epi} sequences used in functional \gls{mri}, and
relies on the acquisition of extra \gls{mri} data. Generally, a \gls{gre} is used to 
obtain a map of magnitude and phase of the actual magnetic field inside the
scanner. Based on this \emph{fieldmap}, a number of methodologies have 
been developed to correct for the distortion, and are generically named as 
\emph{\gls{epi}-unwarp} techniques \cite{holland_efficient_2010,
hsu_correction_2009,jezzard_characterization_2005, reber_correction_2005}. 
Unfortunately, the availability of the corresponding fieldmap is not always
met.\todo[inline]{Mention other problems: low precission, unreliability}

Some other methodologies do not make use of the fieldmaps, compensating the 
distortion with non-linear registration from structural \gls{mri} 
\citep{kybic_unwarping_2000} (see \autoref{sec:registration}), or other means 
\citep{andersson_modeling_2001}. To our knowledge, there exists no study on 
the impact of the \gls{epi} distortion on the variability of tractography 
results.

\subsection{Structural information co-registration}
\label{sec:registration}

@@

\subsection{Summary}
\label{sec:contributions}
